Parallelizing Linear Recurrent Neural Nets Over Sequence Length

### 2017-10-27 paper/main.tex
Added more to synthetic example section

\begin{figure}[t]
\centering
\includegraphics[width=12cm]{cudnn_heatmap.png}
\caption{Throughput comparison between LSTM-256-256 with GILR-LSTM-256-256, with a
  32-dimensional input and a 2-dimensional output, for various batch sizes and
  sequence lengths.The LSTM only has a single row of data because its throughput
  is independent of sequence length. Entries are missing from the GILR-LSTM table
  because there was not enough memory on the GPU to handle such large batch
  sizes and sequences.}
\label{fig:tp_perf}
\end{figure}

@@

Initial experiments showed that the two-layer LSTM was much quicker to converge
than the one-layer LSTM, so here we compare the GILR-LSTM to the two-layer
CudnnLSTM.  We ran all experiments on a NVIDIA K80 GPU, choosing the largest
minibatch size which fit into the GPU memory, with five runs per configuration
allowing us to find the average and standard deviation of the time and number of
iterations to convergence. For all CUDA runs, a brief search over learning rate
and batch size was carried out to find the parameters which allow the network to
converge most rapidly. The criterion for convergence was five consecutive
minibatches giving 100\% accuracy. As can be seen from the learning curves
in figure \ref{fig:synthetic_training}, this was a reasonable criterion. For the
longest sequence-length, we didn't observe the CuDNN-LSTM converging.
\begin{figure}
  \centering
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=1.0\textwidth]{./1k_synthetic.png}
  \end{minipage}%
    \begin{minipage}{0.5\textwidth}
    \includegraphics[width=1.0\textwidth]{./8k_synthetic.png}      
  \end{minipage}
  \begin{minipage}{0.6\textwidth}
    \includegraphics[width=1.0\textwidth]{./1M_synthetic.png}      
  \end{minipage}
  \caption{Training curves for GILR-LSTM and CUDNNLSTM architectures for various
  sequence lengths. (Clockwise from top-left: 1024, 8192, 1048576)}
    \label{fig:synthetic_training}
  \end{figure}

@@

\subsection{Medical Example}
A common application of RNNs is in medicine. In order to illustrate the
increased performance that the plr algorithm allows, we implement an GILR-LSTM
on the problem presented in \Chris{CITE}, the 2016 PhysioNet/Computing in
Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
Heart Sound Recordings challenge is to determine which of a set of EEGs are
produced by people with abnormal hearts.  The data is recorded from a variety of
sources and split into a training and testing set. The testing set is not
publicly available, but can be evaluated through a web portal. The top results
cover a large range of techniques, but generally rely on featurization of the
raw EEG records. We want to see if we are able to leverage our advantage at high
sequence-lenghts to train directly on the raw EEG. The EEGs are sampled at 2,000
Hz and range in length from several seconds to more than one hundred seconds,
giving a max sequence length of over 200,000. Since the testing server is not
equipped with a GPU, we cannot run PLR in the testing step. We resample the
records to 250 Hz so that we are able to evaluate the test results in a
reasonable amount of time, padding all records that are shorter than the longest.

@@

NEED TO GET THOSE RESULTS!

\section{Conclusion}
Parallel linear recurrence is an extremely powerful algorithm and the GILR-LSTM is just one
of many possible models that can be built with it. Future applications of parallel linear recurrence could include sequences orders of magnitude longer than N-MNIST, the development of parallel computable differentiable memory modules, and the combination of linear recurrence with convolutional sequence models. Besides future research, existing models such as Quasi-RNNs and strongly typed RNNs that already contain linear recurrences can immediately benefit from parallel linear recurrence. This work demonstrates the GILR-LSTM significantly accelerates the training of small to medium sized LSTMs. Although similar techniques have been used before, the now explicit concept of linear surrogacy provides a framework for future development and analysis of fast sequence models.

>>>>

\begin{table}[t]
\centering
\includegraphics[width=12cm]{cudnn_heatmap.png}
\caption{Throughput comparison between LSTM-256-256 with GILR-LSTM-256-256, with a
  32-dimensional input and a 2-dimensional output, for various batch sizes and
  sequence lengths.The LSTM only has a single row of data because its throughput
  is independent of sequence length. Entries are missing from the GILR-LSTM table
  because there was not enough memory on the GPU to handle such large batch
  sizes and sequences.}
\label{fig:tp_perf}
\end{table}

@@

Initial experiments showed that the two-layer LSTM was much quicker to converge
than the one-layer LSTM, so here we compare the GILR-LSTM to the two-layer
CudnnLSTM.  We ran all experiments on a NVIDIA K80 GPU, choosing the largest
minibatch size which fit into the GPU memory, with five runs per configuration
allowing us to find the average and standard deviation of the time and number of
iterations to convergence. For all CUDA runs, a brief search over learning rate
and batch size was carried out to find the parameters which allow the network to
converge most rapidly. The criterion for convergence was five consecutive
minibatches giving 100\% accuracy. As can be seen from the learning curves
in figure \ref{fig:synthetic_training}, this was a reasonable criterion. For the
longest sequence-length, we didn't observe the CuDNN-LSTM converging.

The results illustrate that the GILR-LSTM is able to robustly outperform the
CuDNN-LSTM, which is currently the fastest LSTM implementation available. This
is somewhat surprising, as we might expect that the LSTM would do particularly
well at an example that was specifically constructed for it. We can conclude that
the nonlinearities in the LSTM's design are not vital for its performance given
that the linear GILR-LSTM is able to converge, and that if these nonlinearities are
not completely necessary, they entail a large performance cost by diallowing the
plr algorithm from being used.

\begin{figure}
  \centering
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=1.0\textwidth]{./1k_synthetic.png}
  \end{minipage}%
    \begin{minipage}{0.5\textwidth}
    \includegraphics[width=1.0\textwidth]{./8k_synthetic.png}      
  \end{minipage}
  \begin{minipage}{0.6\textwidth}
    \includegraphics[width=1.0\textwidth]{./1M_synthetic.png}      
  \end{minipage}
  \caption{Training curves for GILR-LSTM and CUDNNLSTM architectures for various
  sequence lengths. (Clockwise from top-left: 1024, 8192, 1048576)}
    \label{fig:synthetic_training}
  \end{figure}

@@

\subsection{The Medical Setting}
A common application of RNNs is in medicine. In order to illustrate the
increased performance that the plr algorithm allows, we implement an GILR-LSTM
on the problem presented in \Chris{CITE}, the 2016 PhysioNet/Computing in
Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
Heart Sound Recordings challenge is to determine which of a set of EEGs are
produced by people with abnormal hearts.  The data is recorded from a variety of
sources and split into a training and testing set. The testing set is not
publicly available, but can be evaluated through a web portal. The top results
cover a large range of techniques, but generally rely on featurization of the
raw EEG records. We want to see if we are able to leverage our advantage at high
sequence-lenghts to train directly on the raw EEG. The EEGs are sampled at 2,000
Hz and range in length from several seconds to more than one hundred seconds,
giving a max sequence length of over 200,000. Since the testing server is not
equipped with a GPU, we cannot run PLR in the testing step. We resample the
records to 250 Hz so that we are able to evaluate the test results in a
reasonable amount of time, padding all records that are shorter than the longest.

@@

NEED TO GET THOSE RESULTS!\Chris{Quite likely that the results won't be available
  for tomorrow, or that they won't be particularly impressive. We may have to
  scratch this entire section unfortunately}

\section{Conclusion}
Parallel linear recurrence is an extremely powerful algorithm and the GILR-LSTM
is just one of many possible models that can be built with it. While we were
previously forced to use large minibatches in order to obtain reasonable
training speed, with PLR in many cases we can train rapidly on small minibatch
sizes, freeing us to utilise extremely long sequence lengths and more complex
models. Our CUDA kernel is up to 40 times faster than a serial kernel on
sequence lengths of 65,536, and this benefit transfers to implemented models,
speeding up the SRU \cite{lei2017} by a factor of 9. The impressive results in
\cite{lei2017} and other recent work comprehensively show that
linearly-recurrent RNNs are, in many respects, equally as powerful as nonlinear
LSTMs. Our results on the synthetic example continue this trend, showing that
the GILR-LSTM is able to handle sequence length regimes that are longer than any
we have seen seriously suggested in the literature. We feel that the framework
of describing linear RNNs as linear surrogates for nonlinear RNNs is a fruitful
approach, and that this, paired with the PLR algorithm, will substantially
advance the ability to train RNNs, and to understand timeseries data.

%Future work can the development of parallel computable
%differentiable memory modules, and the combination of linear recurrence with
%convolutional sequence models.

% Besides future research, existing models such as
% Quasi-RNNs and strongly typed RNNs that already contain linear recurrences can
% immediately benefit from parallel linear recurrence. This work demonstrates the
% GILR-LSTM significantly accelerates the training of small to medium sized
% LSTMs. Although similar techniques have been used before, the now explicit
% concept of linear surrogacy provides a framework for future development and
% analysis of fast sequence models.
